Improving \textsl{Gaia} parallaxes with a data-driven model \\ of the color--magnitude diagram

### 2017-06-02 paper/paper.tex
assumptions

To infer the true parallax to a star using the noise-deconvolved \cmd\ as a prior, we make the following assumptions. The method we present is correct
and justifiable under these assumptions, each of which is questionable
in its own right. We return to criticize these assumptions---and the
method that flows from them---in the discussion section below.
\begin{description}
\item[stationarity] We make an assumption of \emph{stationarity}; that is, that the
  stars measured at lower signal-to-noise have analogs measured at
  higher signal-to-noise.
  In detail, this assumption is wrong, (for example, distant halo stars are different than local
  disk stars), but the stationarity assumption is fairly weak:
  The requirement is that, locally in the \cmd, there is
  support among the high signal-to-noise stars for the types
  of stars also seen at low signal-to-noise.
  If this requirement is met, the results will not be
  strongly biased. Although this is a weak assumption, it is deep and fundamental to this
  work.
\item[big data] We assume that we have large numbers of stars, large enough that
  empirical Bayes, or maximum-marginalized likelihood, is a safe
  approximation to full Bayesian inference. This assumption will be
  least true in the least-populated parts of the \cmd. In the extreme
  case, if a portion of the \cmd\ has only one (statistically distinct)
  exemplar, for example, a white dwarf, inference for these cases will be biased.
\item[noise model] We assume that the \tgas\ parallax uncertainties dominate the
  noise in any estimates of absolute magnitude. We further assume that
  the parallax and color uncertainties are Gaussian in form with
  correctly known variances. In detail, we assume that the parallax uncertainty estimates, the parallax{\_}error entry in the \gaia\ catalogue, are correct.
  Here we are making use of the fact that, for
  de-noising, under-estimation of measurement noise is conservative, and leads to under-deconvolution. In other words, the features of the inferred \cmd\ will typically be broader and more conservatively estimated than than if we over-estimated the noise.
\item[dust] We treat the \panstarrs-based three-dimensional dust maps (\citealt{green15})
  as correct in their median dust estimates, and their
  effects on color and magnitude. As we optimize the empirical Bayes model, we iteratively update each star's
  dust correction at its updated, inferred distance. Our assumption is that these corrections are
  good enough, and uncertainties in these are not dominant, for neither
  color nor absolute magnitude.
\item[mixture of Gaussians] We assume that the \cmd\ can be represented by a mixture of
  Gaussians in a particular transformed space. This mixture we fix
  (with only heuristic analysis) at 128 Gaussians.
\item[no physics] We make no use of stellar models; our only assumptions about
  stellar physics are generic and implicit (for example, that the
  \cmd\ is somehow smooth and compact).
\end{description}

\BLcomment{Is it the right place to comment that this whole approach also assumes that 1) the selection function is totally known (and neglected here), 2) all measurements and their errors are correct (unbiased and statistically representative)? This could also go in the discussion, but I personally see those as fundamental assumptions.}

@@

  For very rare kinds of stars. Inference with our prior is bad for
  underrepresented populations in the \cmd. For example, it appears
  there are only $2$ white dwarfs in \gaia\ DR1. With them being very
  underrepresented in the prior, the amplitudes of the gaussians
  representing them is very small, so the WDs are more likely to be
  pulled to mimic a much farther away MS star.
\item[big data] DWH
\item[noise model] DWH
\item[dust] Within our probabilistic framework, we improperly handle
  dust. We are taking a point estimate of the probabilistic dust model
  at a quantile of our distance posterior, correcting our photometry
  using this point estimate, and then re-deriving the prior. A full,
  proper account would sample the dust model at distances sampled from
  the posterior, and propagate those samples back into the prior
  during the iterative process. The reason we don't do a full
  probabilistic treatment, is that this proper dust distribution would
  not be a Gaussian distribution, which \xd\ requires. We could
  calculate this more complex dust distribution and estimate it as a
  Gaussian distribution to feed to \xd, however, this is beyond the
  scope of this paper.
\item[mixture of Gaussians] DWH
\item[no physics] DWH
\end{description}

\BLcomment{it's weird to end with these subsections. Cross-validation should go in the methods part (I think there is a flag there already), and the discussion of assumptions should go in the discussion. There should be a conclusion and look towards the future.}

>>>>

To infer the true parallax to a star using the noise-deconvolved \cmd\ as
a prior, we make the following assumptions. The method we present is correct
and justifiable under these assumptions, each of which is questionable
in its own right. We return to criticize these assumptions---and the
method that flows from them---in the discussion section below.
\begin{description}
\item[stationarity] We make an assumption of \emph{stationarity}; that is, that the
  stars measured at lower signal-to-noise have analogs measured at
  higher signal-to-noise.
  In detail, this assumption is wrong, (for example, distant halo stars are different than local
  disk stars), but the stationarity assumption is fairly weak:
  The requirement is that, locally in the \cmd, there is
  support among the high signal-to-noise stars for the types
  of stars also seen at low signal-to-noise.
  If this requirement is met, the results will not be
  strongly biased. Although this is a weak assumption, it is deep and fundamental to this
  work.
\item[selection] We have no model for the selection function of \tgas,
  nor selection volumes for any kinds of stars.
  For this reason, we are building a model of \emph{the contents of \tgas}, not the
  properties of the stars in the Milky Way (nor any volume-limited component or region
  thereof).
\item[big data] We assume that we have large numbers of stars, large enough that
  empirical Bayes, or maximum-marginalized likelihood, is a safe
  approximation to full Bayesian inference. This assumption will be
  least true in the least-populated parts of the \cmd. In the extreme
  case, if a portion of the \cmd\ has only one (statistically distinct)
  exemplar, for example, a white dwarf, inference for these cases will be biased.
\item[noise model] We assume that the \tgas\ parallax uncertainties dominate the
  noise in any estimates of absolute magnitude. We further assume that
  the parallax and color uncertainties are Gaussian in form with
  correctly known variances. In detail, we assume that the parallax uncertainty estimates,
  (the \texttt{parallax{\_}error} entry in the \gaia\ catalog), are correct.
  Here we are making use of the fact that, for
  de-noising, under-estimation of measurement noise is conservative, and leads to under-deconvolution. In other words, the features of the inferred \cmd\ will typically be broader and more conservatively estimated than than if we over-estimated the noise.
\item[dust] We treat the \panstarrs-based three-dimensional dust maps (\citealt{green15})
  as correct in their median dust estimates, and their
  effects on color and magnitude. As we optimize the empirical Bayes model, we iteratively update each star's
  dust correction at its updated, inferred distance. Our assumption is that these corrections are
  good enough, and uncertainties in these are not dominant, for neither
  color nor absolute magnitude.
\item[mixture of Gaussians] We assume that the \cmd\ can be represented by a mixture of
  Gaussians in a particular transformed space. This mixture we fix
  (with only heuristic analysis) at 128 Gaussians.
\item[no physics] We make no use of stellar models; our only assumptions about
  stellar physics are generic and implicit (for example, that the
  \cmd\ is somehow smooth and compact).
\end{description}

@@

  For very rare kinds of stars. Inference with our prior is bad for
  underrepresented populations in the \cmd. For example, it appears
  there are only $2$ white dwarfs in \gaia\ DR1. With them being very
  underrepresented in the prior, the amplitudes of the gaussians
  representing them is very small, so the WDs are more likely to be
  pulled to mimic a much farther away MS star.
\item[selection] DWH
\item[big data] DWH
\item[noise model] DWH
\item[dust] Within our probabilistic framework, we improperly handle
  dust. We are taking a point estimate of the probabilistic dust model
  at a quantile of our distance posterior, correcting our photometry
  using this point estimate, and then re-deriving the prior. A full,
  proper account would sample the dust model at distances sampled from
  the posterior, and propagate those samples back into the prior
  during the iterative process. The reason we don't do a full
  probabilistic treatment, is that this proper dust distribution would
  not be a Gaussian distribution, which \xd\ requires. We could
  calculate this more complex dust distribution and estimate it as a
  Gaussian distribution to feed to \xd, however, this is beyond the
  scope of this paper.
\item[mixture of Gaussians] DWH
\item[no physics] DWH
\end{description}
